Sources: 
https://www.coursera.org/lecture/financial-risk-management-with-r/calculating-daily-returns-jenBp
https://www.coursera.org/lecture/financial-risk-management-with-r/calculating-longer-returns-m1ZnI
```{r}
library(quantmod)
wil<-getSymbols("WILL5000IND",src="FRED",auto.assign=FALSE)
wil<-na.omit(wil)
wil<-wil["1979-12-31/2017-12-31"]
names(wil)<-"TR"
head(wil,3)
tail(wil,3)
logreturn <- diff(log(wil$TR))[-1]
round(head(logreturn, 3),6)


```
daily return : $re$$t_t$$=$$wil_t$$/$$wil_{t-1}$$-1$


A discrete return can never have a value less than -1, but it can take on a positive number greater than 1. This mean that the distrubution of discrete returns cannot have a symetric distrobution, so we should transform by taking the log of the data


$logre$$t_t$$=log(1+$$ret_t$$)$

To get a n-day return simply add up all the daily returns 

```{r}
logreturn_w <- apply.weekly(wil,sum) #also works for monthly, quarterly, and yearly
round(head(logreturn_w,3),6)

mu <- round(mean(logreturn),8)
sigma <- round(sd(logreturn),8)
mu 
sigma
```

Value at rsik : the ammount that the portfolio might lose, with a given probability (1 - alpha) over a given time period (alpha is usually 0.1, o.05 or 0.01) and the time period is usually one day or one week

